---
title: "Lab 12 Matlab Conversion"
author: "Aaron Swoboda and Oliver Hall"
date: "2025-04-26"
output: pdf_document
---

This file aims to convert a series of Matlab scripts and data files into a R Markdown file that replicates the Tol (2023) Lab 12 for the economics of climate change. https://github.com/rtol/ClimateEconomics 

The Matlab files this replicates are:

* MLIAM.mat
* init01.m
* MRHparam.m
* MRH.m
* RFparam.m
* RadForc.m
* STparam.m
* ST.m
* init02.m
* Popparam
* Outputparam
* CO2param
* init03.m
* CobbDouglas.m
* invest.m
* init04.m
* init05.m
* init08.m
* impactcc.m
* impactcca.m
* init12.m
* MLIAM012.m
* fMLIAM012.m

## Overview - the MLIAM11.m script
```{r, eval=FALSE}
# clear all
# 
# ctax0 = 0;
# citer0 = [0 0 0];
# 
# NPV0 = -fMLIAM12(0.0, citer0, 1);
# 
# options  =  optimset('MaxFunEvals',1000,'MaxIter',1000); 
# 
# ctax1 = fminsearch(@(ctax) fMLIAM11(ctax,citer0,1), ctax0, options);
# ctax2 = fminsearch(@(ctax) fMLIAM11(ctax,citer0,2), ctax0, options);
# ctax3 = fminsearch(@(ctax) fMLIAM11(ctax,citer0,3), ctax0, options);
# 
# citer1(1) = ctax1;
# citer1(2) = ctax2;
# citer1(3) = ctax3;
# 
# ctax1 = fminsearch(@(ctax) fMLIAM11(ctax,citer1,1), ctax0, options);
# ctax2 = fminsearch(@(ctax) fMLIAM11(ctax,citer1,2), ctax0, options);
# ctax3 = fminsearch(@(ctax) fMLIAM11(ctax,citer1,3), ctax0, options);
# 
# citer2(1) = ctax1;
# citer2(2) = ctax2;
# citer2(3) = ctax3;
# 
# ctax1 = fminsearch(@(ctax) fMLIAM11(ctax,citer2,1), ctax0, options);
# ctax2 = fminsearch(@(ctax) fMLIAM11(ctax,citer2,2), ctax0, options);
# ctax3 = fminsearch(@(ctax) fMLIAM11(ctax,citer2,3), ctax0, options);
# 
# citer3(1) = ctax1;
# citer3(2) = ctax2;
# citer3(3) = ctax3;
```

## Step 1 - load the MLIAM.mat file
```{r}
library(R.matlab)
#MLIAM <- readMat("../TolMatlabFiles/MLIAM.mat")
MLIAM <- readMat("./MLIAM.mat")
ls.str()

CO2global <- MLIAM$CO2global        # should we use this?
CO2emit <- MLIAM$CO2emit
# CO2concobs  <- MLIAM$CO2concobs
population  <- MLIAM$population
# output  <- MLIAM$output
# atmtempobs  <- MLIAM$atmtempobs
# oceantempobs  <- MLIAM$oceantempobs
# CO2emitobs  <- MLIAM$CO2emitobs    # should we use this?
outputobs  <- MLIAM$outputobs
energy  <- MLIAM$energy
NReg  <- MLIAM$NReg
```

## Step 2 - run necessary scripts from previous labs and define the usual functions
```{r}
## init01
climsens = 4.260547;
EndYear = 2300;
StartYear = 1750;
year = StartYear:EndYear
NYear = length(year)

# %MRHparam
MRHlife <- c(0, 1-exp(-1/363), 1-exp(-1/74), 1-exp(-1/17), 1-exp(-1/2))
# 
MRH1750 <- c(275, 0, 0, 0, 0);
# 
MRHshare <- c(0.13, 0.20, 0.32, 0.25, 0.10);
# 
CO2convert = 1/2.13/1000;
# 
MRHbox = matrix(0, NYear, 5); # "five boxes" for each year
MRHbox[1, ] = MRH1750;
CO2conc = matrix(0, NYear, 1);
CO20 = sum(MRH1750);
CO2conc[1] = CO20;

# %RFparam
### RadForc
RadForc <-function(CO2,RF0,CO20) RF0*log(CO2/CO20);
# 
RF0 = 5.35;
RF = matrix(0, NYear, 1);
RF[1] = RadForc(CO20, RF0, CO20) #Needs the RadForc function

# %STparam
atmtemp0 = 0;
oceantemp0 = 0;

atmtemp = matrix(0, NYear, 1);
oceantemp = matrix(0, NYear, 1);

atmtemp[(1)] = atmtemp0;
oceantemp[(1)] = oceantemp0;

STpar <- c(climsens/RF0/log(2), 0.0256, 0.00738, 0.00568)

## init02
consumption <- matrix(0, nrow = NYear, ncol = NReg)

# %Popparam
popgrowth <- population[2010 - StartYear + 1, ] / population[2009 - StartYear + 1, ] - 1
popdecline <- 0.95
popgrowth <- popdecline * popgrowth

# %Outputparam
capital <- matrix(0, nrow = NYear, ncol = NReg)
TFP <- matrix(0, nrow = NYear, ncol = NReg)
output <- matrix(0, nrow = NYear, ncol = NReg)
outputpc <- matrix(0, nrow = NYear, ncol = NReg)

TFPgrowth <- c(0.0206, 0.0260, 0.0236)
TFPdecline <- 0.99
outputpc2010 <- c(33498, 3170, 954)

lambda <- 0.2
savings <- 0.2
depreciation <- 0.1

A0 <- (outputobs[1960 - StartYear + 1, ] / population[1960 - StartYear + 1, ])^(1 - lambda)
A0 <- A0 * (depreciation / savings)^lambda
K0 <- (A0 ^ (1 / (1 - lambda))) * population[1960 - StartYear + 1, ]
K0 <- (savings / depreciation)^(1 / (1 - lambda)) * K0

capital[1960 - StartYear + 1, ] <- K0
TFP[1960 - StartYear + 1, ] <- A0

# %CO2param
CO2emitbau <- CO2emit
energyint <- energy / outputobs
emissint <- CO2emit / energy

AEEI <- energyint[(2010 - StartYear + 1), ] / energyint[(1960 - StartYear + 1), ]
AEEI <- AEEI ^ (1 / 50)
ACEI <- emissint[(2010 - StartYear + 1), ] / emissint[(1960 - StartYear + 1), ]
ACEI <- ACEI ^ (1 / 50)

AEEI <- c(0.98926539, 0.98782002, 0.99025746)
ACEI <- c(0.99594960, 1.00029674, 1.00979371)

### MRH - Maier-Reimer Hasselmann model
MRH <- function(CO2concold, CO2emit, CO2life, CO2share, CO2convert) {
  CO2concnew = (1-CO2life)*CO2concold + CO2convert*CO2share*CO2emit
  CO2concnew
}

### Schneider-Thompson model of Ocean and Atm Temps
# updates the temperature of the atmosphere and the ocean using the Schneider-Thompson model
ST <- function(atmtempold,oceantempold,radforc,STpar) {
atmtempnew   = atmtempold   + 
    STpar[2]*(STpar[1]*radforc-atmtempold) + 
    STpar[3]*(oceantempold-atmtempold)
oceantempnew = oceantempold + 
  STpar[4]*(atmtempold-oceantempold)
temps <- c(atmtempnew, oceantempnew)
names(temps) <- c("atm", "ocean")
temps
}

### Cobb-Douglas function
CobbDouglas <- function(A, K, L, lambda) {
  Y <- A * (K^lambda) * (L^(1 - lambda))
  return(Y)
}

### invest function
invest <- function(oldK, Y, s, d) {
  newK <- (1 - d) * oldK + s * Y
  return(newK)
}

## init03
unitabcost <- c(0.1, 0.1, 0.1)
emred <- matrix(0, nrow = NYear, ncol = NReg)
relabcost <- matrix(0, nrow = NYear, ncol = NReg)
totabcost <- matrix(0, nrow = NYear, ncol = NReg)
margabcost <- matrix(0, nrow = NYear, ncol = NReg)

PRTP <- 0.03
RRA <- 1

discountrate <- matrix(0, nrow = NYear, ncol = NReg)
globalDR <- matrix(0, nrow = NYear, ncol = 1)

## init04
emitalloc <- matrix(0, nrow = NYear, ncol = NReg)
emitalloc[266:NYear, ] <- emitalloc[266:NYear, ] + 0.05
permittrade <- matrix(0, nrow = NYear, ncol = NReg)
permitprice <- matrix(0, nrow = NYear, ncol = 1)

## init05
partol <- matrix(c(5.88, -2.31, 0, 3.57, -1.70, 0, 1.96, -1.26, 0), nrow = 3, ncol = 3)
parweitzman <- matrix(c(0, 0.5563, -0.0113, 0, 0.2561, -0.0106, 0, 0.0655, -0.0101), nrow = 3, ncol = 3)
impactpar <- partol
impelas <- 0
impact <- matrix(0, nrow = NYear, ncol = NReg)

## init08
perturbation <- rep(0, NYear)
StartPolicy <- 2015
NPol <- 10
StepPol <- 10

welfare <- matrix(0, nrow = NYear, ncol = NReg)
utilDF <- rep(0, NYear)
utilDF[StartPolicy - StartYear + 1] <- 1
for (i in 2:(EndYear - StartPolicy + 1)) {
  utilDF[StartPolicy - StartYear + i] <- utilDF[StartPolicy - StartYear + i - 1] / (1 + PRTP)
}
### impactcc function
impactcc <- function(temp, impar, impelas, inccap, inccap0) {
  imp <- impar[1,] * temp + impar[2,] * temp^2 + impar[3,] * temp^6
  imp <- imp * (inccap / inccap0)^impelas
  return(imp)
}
  
### impactcca function
impactcca <- function(temp, impar) {
  adap <- -0.5 * temp * impar[1, ] / impar[2, ]
  impa <- impar[1, ] * temp * (1 - adap) - impar[2, ] * adap^2
  imp <- impar[3, ] * temp + impar[4, ] * temp^2
  
  return(list(adap = adap, impa = impa, imp = imp))
}

```

##3 Step 3 - run init12
```{r}
adaptpar = matrix(c(-5.46, -5.05, -4.43, 59.5, 36.4, 32.0, -5.46, -5.05, -4.43, 0.13, 0.17, 0.15), nrow = 4, ncol = 3, byrow = TRUE)

impacta <- matrix(0, nrow = NYear, ncol = NReg)
adaptation <- matrix(0, nrow = NYear, ncol = NReg)
```

### Step 5 - define fMLIAM12 function
```{r}
fMLIAM12 <- function(ctax, coth, reg) {
  if (ctax < 0) {
    NPV <- 1e+15
  } else {
    NPV <- 0
    
    carbontax <- coth
    carbontax[reg] <- ctax
    
    for (t in (StartYear + 1):EndYear) {
      i <- t - StartYear + 1
      MRHbox[i, ] <- MRH(MRHbox[i - 1, ], CO2global[i - 1], MRHlife, MRHshare, CO2convert)
      CO2conc[i] <- sum(MRHbox[i, ])
      RF[i] <- RadForc(CO2conc[i], RF0, CO20)
      temps <- ST(atmtemp[i - 1], oceantemp[i - 1], RF[i], STpar)
      atmtemp[i] <- temps[1]
      oceantemp[i] <- temps[2]
      impact[i, ] <- impactcc(atmtemp[i], impactpar, 0, 1, 1)
      
      if (t > 2010) {
        population[i, ] <- population[i - 1, ] * (1 + popgrowth)
        popgrowth <- popdecline * popgrowth
        TFPgrowth <- TFPdecline * TFPgrowth
      }
      
      if (t > 1960) {
        capital[i, ] <- invest(capital[i - 1, ], output[i - 1, ], savings, depreciation)
        TFP[i, ] <- TFP[i - 1, ] * (1 + TFPgrowth)
      }
      
      if (t >= 1960) {
        output[i, ] <- pmax(0.3 * population[i, ], CobbDouglas(TFP[i, ], capital[i, ], population[i, ], lambda) * (1 - relabcost[i - 1, ] + impacta[i - 1, ] / 100)) # explicit adaptation
        # output[i, ] <- pmax(0.3 * population[i, ], CobbDouglas(TFP[i, ], capital[i, ], population[i, ], lambda) * (1 - relabcost[i - 1, ] + impact[i - 1, ] / 100)) # implicit adaptation
        outputpc[i, ] <- output[i, ] / population[i, ] * 1000
        consumption[i, ] <- (1 - savings) * output[i, ]
        welfare[i, ] <- population[i, ] * log(consumption[i, ])
        # impact[i, ] <- impactcc(atmtemp[i], impactpar, impelas, outputpc[i, ], outputpc2010)
        adaptparams <- impactcca(atmtemp[i], adaptpar)
        adaptation[i, ] <- adaptparams$adap
        impacta[i, ] <- adaptparams$impa
        impact[i, ] <- adaptparams$imp
        
      }
      
      if (t > 2010) {
        energyint[i, ] <- AEEI * energyint[i - 1, ]
        emissint[i, ] <- ACEI * emissint[i - 1, ]
        energy[i, ] <- energyint[i, ] * output[i, ]
        CO2emitbau[i, ] <- emissint[i, ] * energy[i, ]
        
        emred[i, ] <- pmin(0.99, 0.5 * 0.001 * carbontax * CO2emitbau[i, ] / unitabcost / output[i, ])
        
        # permit trade
        # permitprice[i] <- 1000 * emitalloc[i, ] * CO2emitbau[i, ]' / sum(0.5 * CO2emitbau[i, ]^2 / unitabcost / output[i, ])
        # emred[i, ] <- 0.5 * 0.001 * permitprice[i] * CO2emitbau[i, ] / unitabcost / output[i, ]
        # permittrade[i, ] <- (emitalloc[i, ] - emred[i, ]) * CO2emitbau[i, ]
        
        CO2emit[i, ] <- CO2emitbau[i, ] * (1 - emred[i, ])
        CO2global[i] <- CO2global[i - 1] * sum(CO2emit[i, ]) / sum(CO2emit[i - 1, ]) + perturbation[i]
        
        discountrate[i, ] <- PRTP + RRA * (consumption[i, ] / consumption[i - 1, ] - 1)
        globalDR[i] <- PRTP + RRA * (sum(consumption[i, ]) / sum(consumption[i - 1, ]) - 1)
        carbontax <- carbontax * (1 + discountrate[i, ])
      }
      
      if (t > 2015) {
        relabcost[i, ] <- unitabcost * emred[i, ]^2 + 0.001 * permitprice[i] * permittrade[i, ] / output[i, ]
        totabcost[i, ] <- relabcost[i, ] * output[i, ]
        margabcost[i, ] <- 2 * unitabcost * emred[i, ] * output[i, ] / CO2emit[i, ] * 1000
      }
    }
    
    NPV <- -t(utilDF) %*% welfare[, reg]
    
    temp <- atmtemp[450]
  }
  
  return(NPV)
}
```

### Step 6 - run MLIAM12
```{r}
ctax0 <- 0
citer0 <- c(0, 0, 0)

NPV0 <- -fMLIAM12(0.0, citer0, 1)

options <- list(maxit = 1000, reltol = 1e-8)

ctax1 <- optimize(function(ctax) -fMLIAM12(ctax, citer0, 1), c(0, 1), maximum = FALSE)$minimum
ctax2 <- optimize(function(ctax) -fMLIAM12(ctax, citer0, 2), c(0, 1), maximum = FALSE)$minimum
ctax3 <- optimize(function(ctax) -fMLIAM12(ctax, citer0, 3), c(0, 1), maximum = FALSE)$minimum

citer1 <- c(ctax1, ctax2, ctax3)

ctax1 <- optimize(function(ctax) -fMLIAM12(ctax, citer1, 1), c(0, 1), maximum = FALSE)$minimum
ctax2 <- optimize(function(ctax) -fMLIAM12(ctax, citer1, 2), c(0, 1), maximum = FALSE)$minimum
ctax3 <- optimize(function(ctax) -fMLIAM12(ctax, citer1, 3), c(0, 1), maximum = FALSE)$minimum

citer2 <- c(ctax1, ctax2, ctax3)

ctax1 <- optimize(function(ctax) -fMLIAM12(ctax, citer2, 1), c(0, 1), maximum = FALSE)$minimum
ctax2 <- optimize(function(ctax) -fMLIAM12(ctax, citer2, 2), c(0, 1), maximum = FALSE)$minimum
ctax3 <- optimize(function(ctax) -fMLIAM12(ctax, citer2, 3), c(0, 1), maximum = FALSE)$minimum

citer3 <- c(ctax1, ctax2, ctax3)
```